\LARGE \bf ***

### 2020-09-09 icra-2021.tex
friday's countdown (aka preparing a stable version for internal review [1])

@@

    booktitle={Proceedings of the 2020 Fourth IEEE International Conference on Robotic Computing (IRC)},

@@

%% author names and affiliations
\author{
  Adam Seewald$^{*}$, Hector Garcia de Marina, and Ulrik Pagh Schultz
  \thanks{$^{*}$Corresponding author; email: {\tt\small ads@mmmi.sdu.dk}}
  \thanks{Adam Seewald, Ulrik Pagh Schultz are with the SDU UAS Center, M{\ae}rsk Mc-Kinney M{\o}ller Institute, University of Southern Denmark, Odense, Denmark.}
  \thanks{Hector Garcia de Marina is with the Faculty of Physics, Department of Computer Architecture and Automatic Control, Universidad Computense de Madrid, Spain.}
}

@@

Mobile robots are increasingly used in a broad range of applications. The growing demand for autonomy of such devices challenges traditional methodologies, soliciting the development of new and unprecedented techniques. The presented approach proposes a \emph{mission-aware energy planning algorithm} that addresses the broader problem of increasing \emph{computational demands} and their relation to the robot's energy consumption and autonomy. The energy is varied by means of mission-specific parameters, the Quality of Service (QoS) of the onboard computations, and the trajectory explicit equations (TEEs). The approach ideally accounts for a mission extension by adapting the TEEs and QoS as the robot batteries drain.

@@

Given a position $\mathbf{p}\in\mathbb{R}^3$ in the 3D Euler space, with respect to some inertial navigation frame $\mathcal{O}_W$, a guidance action--which affects the position allowing the motion along the 3-axis--is built upon the direction to follow, derived using vector field~\cite{de2017guidance}.

---

%The motion defines a trajectory and a position $\mathbf{p}$; the trajectory is expressed through explicit trajectory equation $\varphi$, while the position determines a vehicle frame $\mathcal{O}_V$ in the 2D Euler space relative to an inertial frame $\mathcal{O}_W$. For the sake of simplicity, it is assumed that the $z$-axis is constrained to a specific altitude $h$, with the system being free to evolve in the $x,y$-axes (i.e., the drone is flying, performing a mission, at a given height). A motion guidance action is derived using vector field design, enabling the convergence to the desired trajectory anywhere in $\mathcal{O}_W$.

%The energy $\mathbf{q}$ spent to perform such guidance action is later derived using Fourier analysis in its state-space representation. This component is decomposed in the energy needed for the actuation (i.e., the mechanical energy), and the computations (i.e., the computational energy), an approach that has been extensively reviewed in our previous work~\cite{seewald2019coarse, seewald2019component, seewald2020mechanical}. A control action, which varies quality of service (QoS) levels of the software being executed, is derived for this purpose. The main goal is to tune the explicit energy equation allowing to maximize the computational and mechanical outcome of the mission while trying to minimize the eventuality of battery discharge.

\subsection{Motion and Guidance Model}
The position $\mathbf{p}$, derived from~\cite{de2017guidance}, is described using the following non-holonomic model
\begin{equation}\label{eqq:1}\begin{cases}
  \dot{\mathbf{p}}(t)&=s\Psi(\psi(t))+w(t)\\
  \dot{\psi}(t)&=u(t)
\end{cases},
\end{equation}
where $\mathbf{p}(t)\in\mathbb{R}^2$ is a point in $\mathcal{O}_W$, $s\in\mathbb{R}$ describes the airspeed assumed constant, $\psi(t)\in(-\pi,\pi]$ the attitude yaw angle and $\Psi(\psi(t))=\begin{bmatrix}\cos{\psi(t)} & \sin{\psi(t)}\end{bmatrix}^T$, and $u\in\mathbb{R}$ the guidance action which denotes the angular velocity $\dot{\psi}$ of $\mathcal{O}_V$.

Let us define a generic continuously differentiable function $\varphi:\mathbb{R}^2\rightarrow\mathbb{R}$. The desired trajectory $\mathcal{P}$ can be defined
\begin{equation}\label{eqq:2}
  \mathcal{P}:=\{\mathbf{p}:\|\varphi(\mathbf{p})\|<c+\varepsilon\},
\end{equation}
as the set of points that follows an explicit trajectory equation $\varphi(\mathbf{p})=c$ within a given $\varepsilon\in\mathbb{R}_{\geq 0}$.

Given a set of discrete values $c$, $\mathcal{O}_W$ can be covered by all the trajectories $\mathcal{P}$ within specific boundaries $\underline{c}\leq c\leq \overline{c}$, where $\underline{a},\overline{a}$ returns the upper bound and lower bound limit of $a$ from the mission specification, which is a lookup table.

The concept of a trajectory is used later to design a controller that selects $c$ with the highest energy value under the energy budget constraints. Here we design a guidance action $u$ that allows following such trajectory. The guidance is derived using the vector field approach also presented in~\cite{de2017guidance}. The algorithm heads to $\mathcal{P}$ by the means of minimizing the norm in Equation~(\ref{eqq:2}).

Let us define $\varPhi:=\varphi(\mathbf{p})$. The vector field is defined as the desired velocity vector
\begin{equation}\label{eqq:3}
  \dot{\mathbf{p}}_d(\mathbf{p}):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi\in\mathbb{R}^2$ is defined as the gradient of $\varphi$ at the point $\mathbf{p}$ (i.e., its vector field), $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain which adjusts the speed of convergence of the vector field to the desired trajectory.

@@

Let us further define $\hat{\mathbf{p}}:=\mathbf{p}/\|\mathbf{p}\|$, the desired course heading rate $\dot{\chi_d}$ is computed by sensing the position $\mathbf{p}$, the ground velocity $\dot{\mathbf{p}}$, and is expressed
\begin{equation}\label{eqq:4}\begin{split}
  &\dot{\chi}(\dot{\mathbf{p}},\mathbf{p})=-E\frac{\dot{\mathbf{p}}_d}{\|\dot{\mathbf{p}}_d\|^2}\cdot\\
  &\hspace{2ex}\left( E\hat{\dot{\mathbf{p}}}_d\hat{\dot{\mathbf{p}}}_d^T\left( (E-k_e\varPhi)H(\varPhi)\dot{\mathbf{p}}-k_e\nabla\varPhi^T\dot{\mathbf{p}}\nabla\varPhi \right) \right)^T,
\end{split}
\end{equation} 
where $H(\cdot)$ is defined as the Hessian operator; the physical meaning is that the curvature of the desired trajectory has to be known in order to be tracked.

Under the assumption of the airspeed $s>\|w(t)\|$ for $t>0$ (i.e., the constant airspeed is greater than the norm of the wind), the guidance action can be expressed
\begin{equation}\label{eqq:5}
  u(\dot{\mathbf{p}},\mathbf{p},\psi)=\frac{\|\mathbf{p}\|}{s\cos{\beta}}\left( \dot{\mathbf{\chi}}_d(\dot{\mathbf{p}},\mathbf{p})+k_d\hat{\dot{\mathbf{p}}}^TE\hat{\dot{\mathbf{p}}}_d \right),
\end{equation} 
where $k_d\in\mathbb{R}_{\geq 0}$ is the gain which adjusts the speed of the convergence of $\dot{\mathbf{p}}_d$, $\beta=\cos^{-1}{\left( \hat{\dot{\mathbf{p}}}^T\Psi(\psi) \right)}$ is the sideslip angle, and $\dot{\chi}_d$ is given in Equation~(\ref{eqq:4}).

\subsection{Energy Model}

To evaluate the energy spent performing given actuation and computations we consider a Fourier series $f:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}_{\geq 0}$ of an arbitrary order $r$
\begin{equation}\label{eqq:6}
  f(t)=\sum_{n=0}^{r}{a_n\cos{\frac{nt}{\xi}}+b_n\sin{\frac{nt}{\xi}}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a_n, b_n\in\mathbb{R}$ for $n\in\{0,\dotsc,r\}$ the Fourier series coefficients. The Fourier analysis allows accounting for the periodicity of the mission, with a trajectory $\mathcal{P}$ being reiterated over time.

The non-linear model in Equation~(\ref{eqq:6}) can be expressed using an equivalent time-varying state-space model in the following form
\begin{equation}\label{eqq:7}\begin{cases}
  \dot{\mathbf{q}}(t)&=A\mathbf{q}(t)+B\mathbf{u}(t)+w(t)\\
  y(t)&=C\mathbf{q}(t)+v(t)
\end{cases},\end{equation}
where $y(t)\in\mathbb{R}_{\geq 0}$ is the energy evolution of the system being controlled, $w(t)$ the same process noise of the system in Equation~(\ref{eqq:1}), and $v(t)\in\mathbb{R}_{\geq 0}$ the measurement noise. The control $\mathbf{u}$ along with the input matrix $B$ are defined subsequently. The state $\mathbf{q}$ represents the evolution of the Fourier series coefficients in time, which can be expressed along with the state transition matrix, and the output matrix
\begin{equation}\begin{split}
  \mathbf{q}(t)&=\left[\begin{array}{ccccccc}
    a_0 & \vline & a_1 & b_1 & \cdots & a_r & b_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{ccccc}
    1          & \vline & \mathbf{0} & \cdots & \mathbf{0} \\\hline
    \mathbf{0} & \vline & A_1        & \cdots & \mathbf{0} \\
    \vdots     & \vline & \vdots     & \ddots & \vdots     \\
    \mathbf{0} & \vline & \mathbf{0} & \cdots & A_r 
  \end{array}\right],\,A_n=\begin{bmatrix}0 & \frac{n}{\xi} \\ -\frac{n^2}{\xi^2} & 0\end{bmatrix},\\
  C&=\left[\begin{array}{ccccccc}
    1 & \vline & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}(t)\in\mathbb{R}^j$ given $j:=2r+1$, $A\in\mathbb{R}^{j\times j}$ is the state transmission matrix, and $C\in\mathbb{R}^j$ is the output matrix. Furthermore, the first row and column of the matrix $A$ contain zeros row vectors and column vectors respectively.

Motivated by the fact that the system of interest is sampled discrete-time, and for the sake of simplicity, we consider the discretized version of Equation~(\ref{eqq:1})~and~(\ref{eqq:7})
\begin{subequations}\label{eqq:9}\begin{align}
  &\begin{cases}\label{eqq:9a}
  \mathbf{p}_{k+1}&=s\Psi(\psi_k)+w_k\\
  \psi_{k+1}&=u_k
  \end{cases},\\
  &\begin{cases}\label{eqq:9b}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+w_k\\
  y_k&=C\mathbf{q}_k+v_k
  \end{cases}.
\end{align}\end{subequations}

As the system was observed to behave stochastically, with the process and measurement noise evolving in a normal distribution, a Kalman filter~\cite{stengel1994optimal, simon2006optimal} is employed to predict the state $\hat{\mathbf{q}}$. Such a state, along with the predicted output $\hat{y}$, differs from the model state $\mathbf{q}$ and measured output $y$ in Equation~(\ref{eqq:9b}) due to the presence of uncertainty.

The estimation of the state and the update of the predicted output is done using
\begin{subequations}\label{eqq:11}\begin{align}
  K_k&=(CP_{k+1}^-C^T+R)^{-1}(P_{k+1}^-C^T),\\
  \hat{\mathbf{q}}_{k+1}&=\hat{\mathbf{q}}_{k+1}^-+K_k(y_k^s+y_k^c-C\hat{\mathbf{q}}_{k+1}^-),\label{eqq:11b}\\
  P_{k+1}&=(I-K_kC)P_{k+1}^-,\\
  \hat{y}_k&=C\hat{\mathbf{q}}_{k+1},\label{eqq:11d}
\end{align}
\end{subequations}
where $K_k\in\mathbb{R}^j$ is the gain of the Kalman filter, and $I$ the identity matrix. $y_k^s,y_k^c$ are the energy readings: $y_k^s$ the flight controller sensor, i.e., the energy needed for the actuation, and $y_k^c$ the energy of a given software configuration described in Equation~(\ref{eqq:13}) later in this section. The noise covariance matrices $Q\in\mathbb{R}^{j\times j},R\in\mathbb{R}$ indicates the process noise and sensor noise covariance respectively, and $\hat{y}_k\in\mathbb{R}_{\geq 0}$ is the estimated energy.

\subsection{Computational Model}

A computational energy model is built using \powprof{}\footnote{\label{cross-ref:powprof-link}\url{https://bitbucket.org/adamseew/powprofiler}}, an open-source modeling tool that measures empirically software configurations and builds an energy model accordingly~\cite{seewald2019coarse}. Specifically, the tool builds a multivariate linear interpolation which is accessed online at the hand of a lookup table in the optimal control algorithm. The system is modeled as follows. An existing ROS system composes several computationally expensive ROS nodes, allowing to vary the number of computations changing some node-specific quality of service (QoS) values via ROS parameters. The tool builds the energy model using mission specification which, besides other mission parameters, specifies per each ROS node a QoS range.

@@

\subsection{Control Action}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgment}

This work is supported and partly funded by the European Union's Horizon2020 research and innovation program under grant agreement No. 779882 (TeamPlay).

@@

>>>>

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}{Corollary}
\newtheorem{conj}{Conjecture}[section]
\newtheorem{defn}{Definition}[section]
\newtheorem{exmp}{Example}[section]
\newtheorem{rem}{Remark}

@@

    booktitle={2020 4th IEEE International Conference on Robotic Computing (IRC)},

@@

%% author names and affiliations
\author{
  Adam Seewald$^{1}$, Hector Garcia de Marina$^{2}$, and Ulrik Pagh Schultz$^{1}$
  \thanks{This work is supported and partly funded by the European Union's Horizon2020 research and innovation program under grant agreement No. 779882 (TeamPlay).
  }
  \thanks{$^{1}$Adam Seewald, Ulrik Pagh Schultz are with the SDU UAS Center, M{\ae}rsk Mc-Kinney M{\o}ller Institute, University of Southern Denmark, Odense, Denmark. Email: {\tt\small ads@mmmi.sdu.dk}.}
  \thanks{$^{2}$Hector Garcia de Marina is with the Faculty of Physics, Department of Computer Architecture and Automatic Control, Universidad Computense de Madrid, Spain.}
}

@@

Mobile robots are increasingly used in a broad range of applications. The growing demand for autonomy of such devices challenges traditional methodologies, soliciting the development of new and unprecedented techniques. The presented approach proposes a \emph{mission-aware energy planning algorithm} that addresses the broader problem of increasing \emph{computational demands} and their relation to the robot's energy consumption and autonomy. The energy is varied by means of mission-specific parameters, the Quality of Service (QoS) of the onboard computations, and the trajectory explicit equations (TEEs) parameters. The approach ideally accounts for a mission extension by adapting the QoS and TEEs parameters as the robot batteries drain.

@@

Starting with a position in space $\mathbf{p}\in\mathbb{R}^3$ in the 3D Euler space, with respect to some inertial navigation frame $\mathcal{O}_W$, a guidance action--which allows the motion along the 3-axis--is built upon the direction to follow and derived using vector field~\cite{de2017guidance}. To define the direction to follow, we use generic continuously differentiable functions TEEs $\varphi:\mathbb{R}^3\rightarrow\mathbb{R}$, a mathematical abstraction representing the desired trajectory (being the function satisfied $\varphi(\mathbf{p})\rightarrow 0$ for all the points approaching such trajectory).

The energy evolution $\mathbf{q}\in\mathbb{R}^j$ spent to perform such guidance action is described using Fourier analysis (the meaning of $j$ will be explained to the reader in Subsection~\ref{sec:energy-model}) and is decomposed in the energy due to the trajectory (TEEs), and the computations (QoS)--an approach that has been adapted from authors' earlier work on computational energy analysis~\cite{seewald2019coarse, seewald2019component}, and energy estimation of a fixed-wing craft~\cite{seewald2020mechanical}.

\subsection{Position in space and guidance action}
\label{sec:position}

For exemplification, we consider a non-holonomic 2D model of a fixed-wing craft flying at an assigned altitude $h\in\mathbb{R}_{>0}$. We will show, after the results, the requirement being eased to an arbitrary mobile robot model
\begin{equation}\label{eqq:1}\begin{cases}
  \dot{\mathbf{p}}(t)&=s\Psi(\psi(t))+w(t)\\
  \dot{\psi}(t)&=u(\mathbf{p}(t))
\end{cases},
\end{equation}
where $\mathbf{p}(t)\in\mathbb{R}^2$ (against a more generic $\mathbb{R}^3$ introduced earlier), $s\in\mathbb{R}$ describes the airspeed assumed constant, $\psi(t)\in(-\pi,\pi]$ the attitude yaw angle and $\Psi(\psi(t))=\begin{bmatrix}\cos{\psi(t)} & \sin{\psi(t)}\end{bmatrix}^T$, and $u\in\mathbb{R}$ the guidance action which denotes the angular velocity $\dot{\psi}$ of the craft.

Let us define $\mathcal{P}$ the area which discloses all the possible paths within the boundaries $\underline{c}\in\mathbb{R}_{\leq 0},\overline{c}\in\mathbb{R}_{\geq 0}$
\begin{equation}\label{eqq:2}
  \mathcal{P}:=\{\mathbf{p}:\underline{c}\leq\varphi(\mathbf{p})\leq\overline{c}\},
\end{equation}
where $\underline{a},\overline{a}$ returns the upper bound and lower bound limit of $a$ from the mission specification, which is a lookup table.

The concept of area between the bounds is used later to design a controller that selects $c$ with the highest energy value under the energy budget constraints. Here we design a guidance action $u$ that allows following $\varphi$ in such area by the means of minimizing the norm $\|\varphi(\mathbf{p})\|$.

Let us define $\varPhi:=\varphi(\mathbf{p})$. The direction to follow is expressed as the desired velocity vector
\begin{equation}\label{eqq:3}
  \dot{\mathbf{p}}_d(\mathbf{p}):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi\in\mathbb{R}^2$ is defined as the gradient of $\varphi$ at the point $\mathbf{p}$ (i.e., its vector field), $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain which adjusts the speed of convergence.

@@

Let us further define $\hat{\mathbf{p}}:=\mathbf{p}/\|\mathbf{p}\|$, the desired course heading rate $\dot{\chi_d}$ is computed by sensing the position $\mathbf{p}$, the ground velocity $\dot{\mathbf{p}}$, and is expressed
\begin{equation}\label{eqq:4}\begin{split}
  &\dot{\chi}(\mathbf{p})=-E\frac{\dot{\mathbf{p}}_d}{\|\dot{\mathbf{p}}_d\|^2}\cdot\\
  &\hspace{2ex}\left( E\hat{\dot{\mathbf{p}}}_d\hat{\dot{\mathbf{p}}}_d^T\left( (E-k_e\varPhi)H(\varPhi)\dot{\mathbf{p}}-k_e\nabla\varPhi^T\dot{\mathbf{p}}\nabla\varPhi \right) \right)^T,
\end{split}
\end{equation} 
where $H(\cdot)$ is defined as the Hessian operator; the physical meaning is that the curvature of the desired trajectory has to be known in order to be tracked.

Under the assumption of the airspeed $s>\|w(t)\|$ for $t>0$ (i.e., the constant airspeed is greater than the norm of the wind), the guidance action can be expressed
\begin{equation}\label{eqq:5}
  u(\mathbf{p},\psi)=\frac{\|\mathbf{p}\|}{s\cos{\beta}}\left( \dot{\mathbf{\chi}}_d(\dot{\mathbf{p}},\mathbf{p})+k_d\hat{\dot{\mathbf{p}}}^TE\hat{\dot{\mathbf{p}}}_d \right),
\end{equation} 
where $k_d\in\mathbb{R}_{\geq 0}$ is the gain which adjusts the speed of the convergence of $\dot{\mathbf{p}}_d$, $\beta=\cos^{-1}{\left( \hat{\dot{\mathbf{p}}}^T\Psi(\psi) \right)}$ is the sideslip angle, and $\dot{\chi}_d$ is given in Equation~(\ref{eqq:4}).

\subsection{Energy evolution due to trajectory}
\label{sec:energy-model}

To evaluate the energy spent due to trajectory and computations, let us consider a Fourier series $f:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}_{\geq 0}$ of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$
\begin{equation}\label{eqq:6}
  f(t)=\sum_{n=0}^{r}{a_n\cos{\frac{nt}{\xi}}+b_n\sin{\frac{nt}{\xi}}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a_n, b_n\in\mathbb{R}$ for $n\in\{0,\dotsc,r\}$ the Fourier series coefficients.

The non-linear model in Equation~(\ref{eqq:6}) can be expressed using an equivalent time-varying state-space model in the following form
\begin{equation}\label{eqq:7}\begin{cases}
  \dot{\mathbf{q}}(t)&=A\mathbf{q}(t)+B\mathbf{u}(t)\\%+w(t)\\
  y(t)&=C\mathbf{q}(t)%+v(t)
\end{cases},\end{equation}
where $y(t)\in\mathbb{R}_{\geq 0}$ is the energy evolution of the system being controlled. The control $\mathbf{u}$ along with the input matrix $B$ are defined later in Subsection~\ref{sec:control-action}, the state $\mathbf{q}$ mimics the original Fourier series coefficients, and
\begin{equation}\label{eqq:8}\begin{split}
  \mathbf{q}(t)&=\left[\begin{array}{ccccccc}
    a_0 & \vline & a_1 & b_1 & \cdots & a_r & b_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{ccccc}
    1          & \vline & \mathbf{0} & \cdots & \mathbf{0} \\\hline
    \mathbf{0} & \vline & A_1        & \cdots & \mathbf{0} \\
    \vdots     & \vline & \vdots     & \ddots & \vdots     \\
    \mathbf{0} & \vline & \mathbf{0} & \cdots & A_r 
  \end{array}\right],\,A_n=\begin{bmatrix}0 & \frac{n}{\xi} \\ -\frac{n^2}{\xi^2} & 0\end{bmatrix},\\
  C&=\left[\begin{array}{ccccccc}
    1 & \vline & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}(t)\in\mathbb{R}^j$ given $j:=2r+1$, $A\in\mathbb{R}^{j\times j}$ is the state transmission matrix, and $C\in\mathbb{R}^j$ is the output matrix. Furthermore, the first row and column of the matrix $A$ contain zeros row vectors and column vectors respectively.

Motivated by the fact that the system of interest is sampled discrete-time, and for the sake of simplicity, we consider the discretized version of Equation~(\ref{eqq:7}) and add the uncertainty
\begin{subequations}\label{eqq:9}\begin{align}
  &\begin{cases}\label{eqq:9a}
  \mathbf{p}_{k+1}&=s\Psi(\psi_k)+w_k\\
  \psi_{k+1}&=u_k(p_k)
  \end{cases},\\
  &\begin{cases}\label{eqq:9b}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+w_k\\
  y_k&=C\mathbf{q}_k+v_k
  \end{cases},
\end{align}\end{subequations}
where $w_k\in\mathbb{R}$ accounts for the environment uncertainty, and $v_k\in\mathbb{R}$ for the measurement error.

The instantaneous energy can be obtained from the model~(\ref{eqq:9b}) as a linear combination of the state.

\begin{lem}\label{lem:2.1}
  The magnitude of two arbitrary states $\mathbf{q}_{k,0},\mathbf{q}_{k,1}$ described in~(\ref{eqq:8}) along their evolution~(\ref{eqq:9b}) at time instant $k$ depends on the instantaneous energy consumption. Suppose the two behave ideally (with no environment uncertainty and measurement error)
  \begin{equation}
    \|\mathbf{q}_{k,0}\|\geq\|\mathbf{q}_{k,0}\|\iff y_{k,0}\geq y_{k,1}.
  \end{equation}
\end{lem}
\begin{proof}
  *\\
\end{proof}

\begin{thm}\label{thm:2.2}
  Consider a continuously differentiable function $\varphi_k:\mathbb{R}^3\rightarrow\mathbb{R}$ at a time instant $k\in\mathbb{Z}_{>0}$.
  Assume a mobile robot performs a periodic mission free to move in $\mathcal{P}$ defined in~(\ref{eqq:2}), and likewise Lemma~\ref{lem:2.1}, the model behaves ideally. 
  Then the instantaneous energy is a linear combination of the state
  \begin{equation}
    y_k=C\mathbf{q}_k=\sum_{n=0}^r{a_n},
  \end{equation}
  where $a_n\in\mathbf{q}_k$ are the $r+1$ state's components at $k$ with $r$ being a pre-assigned arbitrary order, and $C$ is described by Equation~(\ref{eqq:8}).
\end{thm}
\begin{proof}
  *\\
\end{proof}

\subsection{Energy evolution due to computations}

A computational energy model is built using \powprof{}\footnote{\label{cross-ref:powprof-link}\url{https://bitbucket.org/adamseew/powprofiler}}, an open-source modeling tool that measures empirically software configurations and builds an energy model, presented in authors' previous work~\cite{seewald2019coarse}. Specifically, the tool builds a multivariate linear interpolation which is accessed online at the hand of a lookup table in the optimal control algorithm. The system is modeled as follows. An existing ROS system composes several computationally expensive ROS nodes, allowing to vary the number of computations changing some node-specific quality of service (QoS) values via ROS parameters. The tool builds the energy model using mission specification which, besides other mission parameters, specifies per each ROS node a QoS range.

@@

\subsection{Control action}
\label{sec:control-action}

@@

\section{Algorithm}

\subsection{State estimation}

As the environment uncertainty and measurement error evolve in a normal distribution, we use a Kalman filter~\cite{stengel1994optimal, simon2006optimal} for the purpose of state estimation.

The estimation of the state and the update of the predicted output is done using
\begin{subequations}\label{eqq:11}\begin{align}
  K_k&=(CP_{k+1}^-C^T+R)^{-1}(P_{k+1}^-C^T),\\
  \hat{\mathbf{q}}_{k+1}&=\hat{\mathbf{q}}_{k+1}^-+K_k(y_k^s+y_k^c-C\hat{\mathbf{q}}_{k+1}^-),\label{eqq:11b}\\
  P_{k+1}&=(I-K_kC)P_{k+1}^-,\\
  \hat{y}_k&=C\hat{\mathbf{q}}_{k+1},\label{eqq:11d}
\end{align}
\end{subequations}
where $K_k\in\mathbb{R}^j$ is the gain of the Kalman filter, and $I$ the identity matrix. $y_k^s,y_k^c$ are the energy readings: $y_k^s$ the flight controller sensor, i.e., the energy needed for the actuation, and $y_k^c$ the energy of a given software configuration described in Equation~(\ref{eqq:13}) later in this section. The noise covariance matrices $Q\in\mathbb{R}^{j\times j},R\in\mathbb{R}$ indicates the uncertainty and measurement error covariance respectively, and $\hat{y}_k\in\mathbb{R}_{\geq 0}$ is the estimated energy.

\subsection{Optimal control action}

\subsection{Deployment algorithm}

@@

@@

\newpage
